Towards Spherical Robots For Mobile Mapping In Human Made Environments\tnoteref{mytitlenote}

### 2021-07-06 Latex/rebuttal.txt

We clarified that the shell can enclose all sensor without requiring a 
number of points of connection as individual enlosings for all sensors
would. Further we added the requirement, that the shell is very 
durable.

> Lines 111-116: Modern kinematic scanning systems collect point cloud
> data with high speed, so also kinematic point clouds are often dense
> enough for robust plane determination. And on the other hand, sparse
> point clouds are not optimal for point-to-point matching due to the
> long distances between adjacent points.

We clarified that also the limited FOV plays a role in not finding 
planes for robust correspondences. 
FIXME @Andreas and @Dorit: I (Anton) disagree can you double check?

I (Andreas) seconds the statement, that the limited FOV plays a role
in not finding planes for robust correspondences.

### 2021-07-07 Latex/sections/conclusion.tex
Clustering added. TODO: results.

However, as section~\ref{sec:experimentalSetup} shows, some outliers remain in the registered point cloud due to the simplistic point-to-plane correspondence model.
This is to be improved in future studies.
One idea to do so is to employ a local clustering technique onto the individual scans before registering them globally. 
Points from the scan with a similar normal get clustered together if their distance is small enough.
This way, a cluster represents a local plane while at the same time storing all the points that belong to that plane.
Point-to-plane correspondences then arise from the intersection of local clusters with global planes.

%
Right now, not all steps in the procedure are autonomous, in particular the parameter tuning. 
In the future, one goal is to increase the autonomy of the system.
One approach is to introduce soft-locks for the optimization dimensions.
I.e., instead of locking some dimensions entirely from being used for optimization, they are weighted based on the dynamics of the system that encode which noise source is more likely.  
%
Currently, we are assembling the sphere of Figure~\ref{fig:robotRender} to test it in real environments.
Unfortunately, this sphere will be too large to fit our floating setup, thus it will be tested in an office like environment.
Furthermore, we will test our rolling and scanning spheres on a rail system to achieve repeatable trajectories.

>>>>

%
Right now, not all steps in the procedure are autonomous, in particular the parameter tuning. 
In the future, one goal is to increase the autonomy of the system.
One approach is to introduce soft-locks for the optimization dimensions.
I.e., instead of locking some dimensions entirely from being used for optimization, they are weighted based on the dynamics of the system that encode which noise source is more likely.

%
The biggest issue with the presented algorithm currently is that the global plane model is established only once as initialization, but never updated afterwards.
Utilizing the local planar clustering (LPC) of the individual line scans for updating the global plane model step by step is currently the primary objective of future studies.
One idea is to calculate the convex hulls of the LPC, and combine them with the global planes sequentially.
Therefore, we expect to increase robustness and autonomy of the registration procedure.

%
Currently, we are assembling the sphere of Figure~\ref{fig:robotRender} to test it in real environments.
Unfortunately, this sphere will be too large to fit our floating setup, thus it will be tested in an office like environment.
Furthermore, we will test our rolling and scanning spheres on a rail system to achieve repeatable trajectories.

### 2021-07-07 Latex/sections/registration.tex
Clustering added. TODO: results.

@@

There are many ways of solving the segmentation problem of assigning each point to a plane. They are often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018} have been successful.
The quality of the preregistration directly influences the quality of the plane detection and therefore the quality of the final registration result. 
We assume small enough errors in pre-registration to allow for plane detection.
Therefore, a simple algorithm is sufficient:
For each globally extracted plane, consider adding a correspondence for all points that are close to that plane.
To do this, we employ two distance models to represent the distance from a point to a plane.

@@

We consider a point as ``close'' to a plane, if the Hesse distance is smaller than some threshold $\epsilon_H$ and the PPD is additionally smaller than some other threshold $\epsilon_P$. 
Depending on the threshold values, some points might have correspondences with multiple planes, if both distances fall below their respective threshold. \todo{@Fabi add actual $\epsilon$ values to the experiments}. 
We then choose to establish a correspondence to the plane with the smallest PPD.
If there are still ambiguities, i.e. multiple PPD's with value zero, a correspondence is established with the plane having the smallest PPD and smallest Hesse distance.

>>>>

There are many ways of solving the segmentation problem of assigning each point to a plane. They are often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018} have been successful.
The quality of the preregistration directly influences the quality of the plane detection and therefore the quality of the final registration result.
We assume small enough errors in pre-registration to allow for plane detection.
We employ two distance models to represent the distance from a point to a plane, and combine them with a local planar clustering (LPC) approach to establish point-to-plane correspondences.

\subsubsection{Distance Models}

@@

@@

\subsubsection{Local Planar Clustering}

The presented algorithm employs LPC on each individual line scan.
The idea is to find planar features utilizing an optimized approximate k nearest neighbour (AKNN) search for normal calculation as in~\cite{aknnbbs}.
Thus, normals are calculated for every point in the respective line scan.
Then, the local normal model is combined with a region growing approach.
The region growing is implemented by storing the k-nearest neighbors in a queue, which represents the iteration order of the clustering algorithm.
A region grows if one of these neighbors has a distance shorter then a treshold $d_{growth}$ and additionally has a similar normal.
The angle between two normalized vectors is
\begin{equation}
\label{eq:smallestalpha}
	\alpha(\vec{n}_1, \vec{n}_2) = \arccos(\vec{n}_1 \cdot \vec{n}_2) \; .
\end{equation}
However, the \textit{smallest} angle between two \textit{normal} vectors is 
\begin{equation}
\label{eq:alphahat}
\hat \alpha(\vec{n}_1, \vec{n}_2) = 
\begin{cases}
	2\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{3}{2} \pi \\
	\alpha(\vec{n}_1, \vec{n}_2) - \pi & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \pi \\
	\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{\pi}{2} 
\end{cases}
\; ,
\end{equation}
because the opposing normals $-\vec{n}_1$ and $-\vec{n}_2$ always have to be considered, since they correspond to the same plane.
We say that two normals are similar if the angle $\hat \alpha$ from eq. \ref{eq:alphahat} is smaller than a threshold $\epsilon_{\alpha}$.

Planar areas like walls are therefore connected in one large cluster, while non-feature parts, i.e. non-planar or detached parts of the scan will have their own, smaller cluster.
You are then left with a rather easy task: filtering the clusters.
The filtering has two parts.
First, to combine clusters that are similar, i.e. have a short distance ($d_{growth}$) to each other and similar mean normals.
Second, to filter out all clusters that contain non-feature points.
Since all the non-feature points belong to small clusters, the presented algorihtm does this by putting a minimum threshold $N_{c,min}$ on the cluster size.
Figure \ref{fig:clustering} shows this concept on a real world dataset.
It was obtained by a floating sphere on a desk in an office environment, as described later in section \ref{sec:experimentalSetup}.

\begin{figure*}
	\includegraphics[width=\linewidth]{images/clustering}
	\caption{(Left) Point normals using the AKNN method with $K = 20$. (Middle) Result of the region growing, before applying the second filter. (Right) Resulting clusters after applying the second filter, using $N_{c,min} = 300$. }
	\label{fig:clustering}
\end{figure*}

Point-to-Plane correspondences are now established.
The distance models (Hesse + PPD) are utilized to check wether a cluster overlaps with any of the global planes, i.e. any point in the cluster has a Hesse distance smaller than some threshold $\epsilon_H$ and a PPD smaller than some additional threshold $\epsilon_P$, to the global plane.
For each overlap, we find the smallest angle between the corresponding cluster normal and the global plane normal, according to eq. \ref{eq:alphahat}.
Finally, a correspondence is established between all points in the cluster and the global plane, if the minimum angle between their normals is smaller than $\epsilon_\alpha$.
